---
title: Investigate the exponential distribution in R and compare it with the Central Limit Theorem
author: "Kishibe Rohan"
date: "July 12,2020"
output: pdf_document
---

## Overview
The purpose of this data analysis is to investigate the exponential distribution in R and compare it to the Central Limit Theorem. For this analysis,lambda will be set to 0.2 and the distribution of averages of 40 exponentials over 1000 simulations will be investigated.


## Simulations
Set the parameters for simulations,namely lamba,exponentials and seed.
```{r,echo=TRUE}

#set seed for reproducibility
set.seed(1331)       
lambda <- 0.2

#number of samples & simulations
n <- 40
simulations <- 1000

#Simulate
sim_exp <- replicate(simulations,rexp(n,lambda))

#Calculate mean
means_exp <- apply(sim_exp,2,mean)
```

## Question 1
Show the sample mean and compare it
to the theoretical mean of the distribution.

```{r}
sim_mean <- mean(means_exp)
sim_mean
```

```{r}
theoretical_mean <- 1/lambda
theoretical_mean
```

## Visualization
```{r}
library(ggplot2)
hist(means_exp,xlab="Mean",main="Distribution of means")
abline(v=sim_mean,col="red")
abline(v=theoretical_mean,col="blue")
```


The analytics mean is 4.982624 and the theoretical mean is 5. The center of distribution of averages of 40 exponentials is very close to the theoretical center of the distribution.

## Question 2
Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

```{r}
#standard deviation & variance of sample
sd_sim <- sd(means_exp)
sd_sim
var_sim <- sd_sim ^ 2
var_sim
```

```{r}
#theoretical standard deviation & variance
sd_theory <- (1/lambda)/sqrt(n)
sd_theory
var_theory <- sd_theory^2
var_theory
```
Standard Deviation of the distribution is 0.7830017 while the theoretical standard deviation is 0.7905694.

The theoretical variance is 0.625 while the actual variance of the distribution is  0.6130916

## Question 3
Show that the distribution is approximately normal.

```{r}
x_plt <- seq(min(means_exp),max(means_exp),
             length=100)
y_plt <- dnorm(x_plt,mean = 1/lambda,sd=(1/lambda/sqrt(n)))
hist(means_exp,breaks = n,prob=T,col="salmon",xlab="means",ylab="density",main="Distribution of means")
lines(x_plt,y_plt,pch = 18,col="black",lty = 2)
```

From the plot we observe that the distribution of sample means is approximately the same as the standard normal distribution,apparent from the bell-shaped nature of the curve.
